---
title: "GGPLOT2 in R"
author: "Yiming Zhang"
date: "2/8/2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(ggrepel)
```

## Outline

- Why ggplot2
- Component of Grammar of Graphic
- Basic Structure
- Construct Plot Layer by Layer
- Scales, Position, Facets and Themes
- An Example

## Why ggplot2

- Grammar of graphic
- Both quick and complex plot
- Easy to "create" plot
- Nice aesthetic settings
- Great docummentation and active mailing list 

Things you cannot do:

- 3-dimensional graphics (rgl)
- Interactive graphics (ggvis, plotly)

## Components of Grammar of Graphic {.smaller}

- **Data** that you want to visualise and a set of aesthetic mappings describing how variables in the data are mapped to aesthetic attributes.

- **Layers** made up of geometric elements and statistical transformation. Geometric objects, **geoms** for short, such as points, lines, polygons, etc. Statistical transformations, **stats** for short, summarise data in many useful ways, such as, histogram and summarising a 2d relationship with a linear model.

- The **scales** map values in the data space to values in an aesthetic space, whether it be colour, or size, or shape.

- A **coordinate system**, coord for short, describes how data coordinates are mapped to the plane of the graphic.

- A **facet** describes how to break up the data into subsets and how to display those subsets as small multiples. 

- A **theme** which controls the finer points of display, like the font size and background colour. 

## Data Frame

- [2017 Airbnb data of New York City](http://insideairbnb.com/get-the-data.html) 

- 43234 observations and 16 variables

- It consists of features of each Airbnb room in NYC such as price, review per month, name of neighbourhood, name of borough, latitude and longitude

- Simple cleansing

```{r, include = FALSE}
Airbnb <- read.csv(url('http://data.insideairbnb.com/united-states/ny/new-york-city/2017-10-02/visualisations/listings.csv'))
Airbnb <- Airbnb[ which(Airbnb$price > 0 & Airbnb$price < 500), ]
Airbnb$reviews_per_month[is.na(Airbnb$reviews_per_month)] = 0
```

## Structure of ggplot

ggplot(data, aes(x = ,y = )) + layers + additional elements

```{r, include = TRUE, echo = TRUE, fig.height = 4, fig.width = 7, fig.align='center'}
ggplot(Airbnb, aes(x = reviews_per_month, y = price))
```

## Add a layer

- Display the data or the statistical summaries of the data

- Mainly use geom_xxx() function

- An alternative way is stat_xxx() function

- A plot must have at least one geom or stat function. there is no upper limit. You can add a layer to a plot using the + operator

- All kinds of geom function and stat function: [ggplot2 cheat sheet](https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf)

## Geom_xxx()

```{r, include = TRUE, echo = TRUE, fig.height = 4, fig.width = 7, fig.align='center', warning=FALSE}
ggplot(Airbnb, aes(price, reviews_per_month)) +
  geom_point(size = 0.1) +
  facet_grid(~room_type)
```

## Aesthetic Mapping

- Describe how variables are mapped to visual properties

- aes()

- Specifying the Aesthetics in the Plot(ggplot()) or in the Layers(geom_xxx() or stat_xxx())

- Aesthetic Mapping can consists of Position (i.e., on the x and y axes), color (“outside” color), fill (“inside” color), shape (of points), linetype and size, etc.

## Aesthetic Mapping

```{r, include = TRUE, echo = TRUE}
ggplot(Airbnb, aes(x = price)) +
  geom_histogram(bins = 40, aes(color = room_type), fill = "grey") 
```

## Aesthetic Mapping

```{r, include = TRUE, echo = TRUE}
ggplot(Airbnb, aes(x = price)) +
  geom_histogram(bins = 40, aes(fill = room_type), color = "grey") 
```

## Construct Plot Layer by Layer 

```{r, include = TRUE, echo = TRUE, fig.height = 3.5, fig.width = 7}
ggplot(Airbnb) +
  geom_violin(aes(neighbourhood_group, price), colour = 'blue')
```

## Construct Plot Layer by Layer 

```{r, include = TRUE, echo = TRUE, fig.height = 3, fig.width = 7}
ggplot(Airbnb) +
  geom_violin(aes(neighbourhood_group, price), colour = 'blue') +
  geom_boxplot(aes(neighbourhood_group, price), width = 0.16, 
               outlier.size = 0, notch = TRUE)
```

## Scales {.smaller}

- Controls the mapping from data to aesthetic

- Include position, color and fill, size, shape and line type

```{r, results='asis', echo=FALSE, warning=FALSE}
library(knitr)
Scale <- c('scale_color_','scale_fill_','scale_size_','scale_shape_','scale_linetype_','scale_x_','scale_y_')
Example <- c('scale_color_gradient','scale_fill_discrete','scale_size_manual','scale_shape_discrete','scale_linetype_manual','scale_x_continuous','scale_y_reverse')
kable(data.frame(Scale, Example), caption = "Available Scales")
```

## Scale Modification Example

```{r, include = TRUE, echo = TRUE, fig.height = 4, fig.width = 7}
w1 <- ggplot(Airbnb) +
  geom_bar(aes(x = neighbourhood_group, fill = room_type))
w1
```

## Scale Modification Example

```{r, include = TRUE, echo = TRUE, fig.height = 4, fig.width = 7}
w1 + scale_x_discrete(name = 'Borough') +
  scale_y_continuous(name = 'Number of rooms',
                     breaks = seq(0,20000,4000)) +
  scale_fill_grey(name = 'Room type')
```

## Theme

- Control appearance of non-data elements

- Titles, tick marks and labels

- Legends appearance

- Overall look

- theme()

## Facet

- Facetting generates small multiples each showing a different subset of the data.

- facet_grid() and facet_wrap()

- facet_grid() is fundamentally 2d, being made up of two independent components.

- facet_wrap() is 1d, but wrapped into 2d to save space

## Facet_wrap()

```{r, include = TRUE, echo = TRUE, fig.height = 4, fig.width = 7}
ggplot(Airbnb, aes(price)) +
  geom_histogram(bins = 40) +
  facet_wrap(~neighbourhood_group)
```

## Facet_grid()

```{r, include = TRUE, echo = TRUE,fig.height = 4, fig.width = 7}
ggplot(Airbnb, aes(price)) +
  geom_histogram(bins = 40) +
  facet_grid(room_type ~ neighbourhood_group)
```

## Putting All Things Together

How to construct this plot?

```{r, include = TRUE, echo = FALSE}
Airbnb_sub1 <- aggregate(Airbnb[ ,c(10,14)], list(Airbnb$neighbourhood, Airbnb$neighbourhood_group, Airbnb$room_type), mean)

neighbourhood_count <- data.frame(table(Airbnb$neighbourhood, Airbnb$neighbourhood_group, Airbnb$room_type))
neighbourhood_count <- neighbourhood_count[which(neighbourhood_count$Freq != 0), ]

Airbnb_sub2 <- merge(Airbnb_sub1, neighbourhood_count, 
      by.x =c( 'Group.1', 'Group.2', 'Group.3'), by.y = c('Var1', 'Var2', 'Var3'), all.x = TRUE)
names(Airbnb_sub2) <- c('neighbourhood', 'neighbourhood_group', 'room_type', 'average_price', 'average_rpm', 'count') 

good_neighbourhood <- subset(Airbnb_sub2, Airbnb_sub2$average_rpm >= 3)
good_neighbourhood <- subset(good_neighbourhood, good_neighbourhood$average_price <=150)

ggplot(Airbnb_sub2, aes(average_price, average_rpm)) +
  geom_vline(aes(xintercept = 150), color = 'grey') +
  geom_hline(aes(yintercept = 3), color = 'grey') +
  geom_point(aes(color = neighbourhood_group, size = count), 
             shape = 1, stroke = 1) +
  geom_text_repel(aes(label = neighbourhood), 
                  data = good_neighbourhood, size = 2.5, force = 8) +
  scale_x_continuous(name = 'Average Price', breaks = seq(0, 450, 50)) +
  scale_y_continuous(name = 'Average Reviews Per Month', breaks = c(1:8)) +
  scale_color_discrete(name = '') +
  scale_size_continuous(name = '') +
  labs(title = 'Neibourhood in NYC Airbnb') +
  theme_minimal() +
  theme(legend.position = 'top',
        legend.text = element_text(size = 8, color = 'gray10'),
        plot.title = element_text(size = 15, face = 'bold', hjust = 0.5 )) +
  facet_wrap(~room_type)
```

## Data Transformation

```{r, include = TRUE, echo = TRUE}
Airbnb_sub1 <- aggregate(Airbnb[ ,c(10,14)], 
  list(Airbnb$neighbourhood, Airbnb$neighbourhood_group,
       Airbnb$room_type), mean)
neighbourhood_count <- data.frame(table(Airbnb$neighbourhood,
                  Airbnb$neighbourhood_group, Airbnb$room_type))
neighbourhood_count <- 
      neighbourhood_count[which(neighbourhood_count$Freq != 0), ]
Airbnb_sub2 <- merge(Airbnb_sub1, neighbourhood_count, 
      by.x =c( 'Group.1', 'Group.2', 'Group.3'), 
      by.y = c('Var1', 'Var2', 'Var3'), all.x = TRUE)
      names(Airbnb_sub2) <- c('neighbourhood', 'neighbourhood_group',
                'room_type', 'average_price', 'average_rpm', 'count') 
good_neighbourhood <- 
  subset(Airbnb_sub2, Airbnb_sub2$average_rpm >= 3)
good_neighbourhood <- 
  subset(good_neighbourhood, good_neighbourhood$average_price <=150)
```

## Basic Scatterplot with three facets

```{r, include = TRUE, echo = TRUE,fig.height = 4, fig.width = 7}
p <- ggplot(Airbnb_sub2, aes(average_price, average_rpm)) +
  geom_point(aes(color = neighbourhood_group, size = count)) +
  facet_wrap(~room_type)
p
```

## What need we do?

- Change the point shape to open circle

- Add horizontal and vertical line

- Add name of those 'good neighbourhoods'

- Title, label axes, legend

- Change theme

## Change the point shape to open circle

```{r, include = TRUE, echo = TRUE,fig.height = 3.5, fig.width = 7}
p1 <- ggplot(Airbnb_sub2, aes(average_price, average_rpm)) +
  geom_point(aes(color = neighbourhood_group, size = count),
             shape = 1, stroke = 1) +
  facet_wrap(~room_type)
p1
```

## Add horizontal and vertical line

```{r, include = TRUE, echo = TRUE,fig.height = 4, fig.width = 7}
p2<- p1 +  geom_vline(aes(xintercept = 150), color = 'grey') +
  geom_hline(aes(yintercept = 3), color = 'grey')
p2
```

## Add names of those 'good neighbourhoods'

```{r, include = TRUE, echo = TRUE,fig.height = 3.5, fig.width = 7}
p3 <- p2 + geom_text(aes(label = neighbourhood), 
           data = good_neighbourhood, size = 2.5)
p3
```

## A better way to add those names

```{r, include = TRUE, echo = TRUE,fig.height = 4, fig.width = 7}
library(ggrepel)
p3 <- p2 + geom_text_repel(aes(label = neighbourhood), 
           data = good_neighbourhood, size = 2.5)
p3
```

## Change theme

```{r, include = TRUE, echo = TRUE}
p4 <- p3 + theme_minimal()
p4
```

## Title, label axes and legend

```{r, include = TRUE, echo = TRUE}
p5 <- p4 +
  scale_x_continuous(name = 'Average Price',
                     breaks = seq(0, 400, 100)) +
  scale_y_continuous(name = 'Average Reviews Per Month', 
                     breaks = c(1:8)) +
  scale_color_discrete(name = '') +
  scale_size_continuous(name = '') +
  labs(title = 'Neibourhood in NYC Airbnb') +
  theme(legend.position = 'top',
        legend.text = element_text(size = 8, color = 'gray10'),
    plot.title = element_text(size = 15, face = 'bold', hjust = 0.5))
```

## Title, label axes and legend

```{r, include = TRUE, echo = TRUE}
p5
```

## Additional Resource

- Wickham, H. (2016). ggplot2: elegant graphics for data analysis. Springer.

- [ggplot2 cheat sheet](https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf)

- [ggplot2 workshop](http://tutorials.iq.harvard.edu/R/Rgraphics/Rgraphics.html#faceting)

- [Top 50 ggplot2 Visualizations](http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html)

## Thank you for your audience!